Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-08-27 documents/ovando-etal-assessing-global-fisheries.Rmd
testing TMB fix on windows

What hapens when we impose informative priors on the population parameters? For example at a bare minimum, we know that carrying capacity must be less than infinity, and we can certainly make inferences based on knowledge from other better studied fisheries of similar species on what a potential upper bound on carrying capacity might be. Once we have generated informative priors on our population parameters, we have implicitly provided priors on the biomass and therefore status of the fishery over its entire history: if our priors on carrying capacity only extend a small fraction beyond the minimum value needed to ensure the existence of the population, our implicit prior is that the population is severely depleted. If our prior is that carrying capacity is orders of magnitude greater than the observed catches, then our implicit prior is that the population is relatively unexploited.

@@

For our case study, expert opinion tells us that the stock is likely between XX and XX% of carrying capacity in the most recent year. While we can use `sraplus` to include this addition information, doing so introduces a nontrivial statistical problem: we have now placed two priors on the same outcome (current stock status), one explicit and one implicit in our growth parameter priors.  The net effect of this is that once these priors are run through `sraplus`, our post-model-pre-data "estimate" of stock status will appear to have "updated", creating an illusion that we have learned something about stock status. Remember though that we have absolutely no data on which to update our prior on stock status beyond the minimum floor needed to ensure that the population exists.

`sraplus` implements an approximate solution to this problem. Rather than using combinations of implied and explicit priors, `sraplus` asks, given an explicit prior on recent stock status, what prior distributions of growth parameters produce our explicit prior on recent stock status, given the catch history and model assumptions. This helps resolve some of the statistical issues related to placing multiple priors on the same parameter, and makes clear where any information on recent stock status is coming from. To return to our case study then, uncorrected, combining our explicit priors on recent stock status with our population parameter priors reveals an implicit prior on recent stock status that spans XX to XX. Using our correction, the implicit prior on recent stock status more or less matches our explicit prior (Fig.XX).

While `sraplus` lets users supply their best estimate of recent stock status for a given fishery, the model also allows users to supply data on the swept area ratio (SAR , @amoroso2018) and/or fisheries management index (FMI, @melnychuk2017) scores for a given fishery. SAR is a measure of the trawling intensity in a given region, and as such for species susceptible to trawl gears can be an effective index of fishing pressure. FMI reflects expert opinion on the research, management, enforcement, and socioeconomic capacity of a given fishery, which together have significant associations with stock status. `sraplus` uses relationships between these SAR and FMI scores and observed F/F~MSY~  and B/B~MSY~ values to generate a posterior predictive distribution of F/F~MSY~ or B/B~MSY~ for the fishery in question, which serves as the priors for `sraplus`.

To continue our case study example, suppose that we no longer believe we have good data on current stock status, but do have a reasonable estimate of swept area ratio. `sraplus` now uses the posterior predictive F/F~MSY~ generated by the SAR value as the prior for recent F/F~MSY~ for the fishery in question. This process allows users to translate something that they can measure, SAR, into something that they perhaps cannot (F/F~MSY~). The model then determines the growth parameter priors that conditional on the catch history and model structure produce the SAR generated F/F~MSY~ prior

pull out example from SAR fits
```{r}
sar <- sraplus::sar %>% 
  filter(stockid %in% unique(ram_data$stockid))

ex <- ram_data %>% 
  filter(stockid == "PLAICIS")

@@

@@

# fit using effort index

>>>>

What happens when we impose informative priors on the population parameters? For example at a bare minimum, we know that carrying capacity must be less than infinity, and we can certainly make inferences based on knowledge from other better studied fisheries of similar species on what a potential upper bound on carrying capacity might be. Once we have generated informative priors on our population parameters, we have implicitly provided priors on the biomass and therefore status of the fishery over its entire history: if our priors on carrying capacity only extend a small fraction beyond the minimum value needed to ensure the existence of the population, our implicit prior is that the population is severely depleted. If our prior is that carrying capacity is orders of magnitude greater than the observed catches, then our implicit prior is that the population is relatively unexploited.

@@

For our case study, expert opinion tells us that the stock is likely between XX and XX% of carrying capacity in the most recent year. While we can use `sraplus` to include this addition information, doing so introduces a nontrivial statistical problem: we have now placed two priors on the same outcome (current stock status), one explicit and one implicit in our growth parameter priors.  The net effect of this is that once these priors are run through `sraplus`, our post-model-pre-data "estimate" of stock status will appear to have been updated, creating an illusion that we have learned something about stock status. However, we have absolutely no data with which to update our prior on stock status beyond the requirement that the population exists.

`sraplus` implements an approximate solution to this problem. Rather than using combinations of implied and explicit priors, `sraplus` asks, given an explicit prior on recent stock status, what prior distributions of population parameters produce our explicit prior on recent stock status, given the other prior distributions, catch history and model assumptions. This helps resolve some of the statistical issues related to placing multiple priors on the same parameter (recent stock status), and makes clear where any information on recent stock status is coming from. To return to our case study then, uncorrected, combining our explicit priors on recent stock status with our population parameter priors reveals an implicit prior on recent stock status that spans XX to XX. Using our correction, the implicit prior on recent stock status more or less matches our explicit prior (Fig.XX).

id = unique(ram_comp_data$stockid)

ex <- ram_data %>% 
  filter(stockid %in% id[[1]])

@@

co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
                                 catch = ex$catch, 
                                 years = ex$year,
                                 terminal_state = 0.25,
                                 terminal_state_cv = 0.25,
                                 b_ref_type = "k")

plot_driors(co_driors)

biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
                                      estimate_shape = TRUE)

co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
                                      estimate_shape = TRUE)

plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit)

While `sraplus` lets users supply their best estimate of recent stock status for a given fishery, the model also allows users to supply data on the swept area ratio (SAR , @amoroso2018) and/or fisheries management index (FMI, @melnychuk2017) scores for a given fishery. SAR is a measure of the trawling intensity in a given region, and as such for species susceptible to trawl gears can be an effective index of fishing pressure. FMI reflects expert opinion on the research, management, enforcement, and socioeconomic capacity of a given fishery, which together have significant associations with stock status. `sraplus` uses relationships between these SAR and FMI scores and observed F/F~MSY~  and B/B~MSY~ values to generate a posterior predictive distribution of F/F~MSY~ or B/B~MSY~ for the fishery in question, which serves as the priors for `sraplus`. This process lets us convert information on stock status present in other available data into a form that aligns with the model structure of `sraplus`.

To continue our case study, suppose that we no longer believe we have good data on current stock status, but do have a reasonable estimate of swept area ratio. `sraplus` can now use the posterior predictive $\hat{F/F~MSY~}$ generated by the SAR value as the prior for recent F/F~MSY~ for the fishery in question. This process allows users to translate something that they can measure, SAR, into something that they perhaps cannot (F/F~MSY~). The model then determines the growth parameter priors that conditional on the catch history and model structure produce the SAR generated $\hat{F/F~MSY~}$ prior distribution.

pull out example from SAR fits
```{r}
sar <- sraplus::sar %>% 
  filter(stockid %in% unique(ram_data$stockid)) %>% 
  arrange(desc(sar)) %>% 
  select(stockid, everything()) %>% 
  slice(10)

ex <- ram_data %>% 
  filter(stockid %in% unique(sar$stockid))

co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
                                 catch = ex$catch, 
                                 years = ex$year,
                                 terminal_state = 0.25,
                                 terminal_state_cv = 0.25,
                                 b_ref_type = "k")

plot_driors(co_driors)

biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
                                      estimate_shape = TRUE)

co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
                                      estimate_shape = TRUE)

plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit)

@@

sar_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
                                 catch = ex$catch, 
                                 years = ex$year,
                                 sar = sar$sar[sar$stockid == unique(ex$stockid)])

plot_driors(sar_driors)

sar_fit <- fit_sraplus(sar_driors, estimate_shape = TRUE)

plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit, `sar fit` = sar_fit)

# fit using biomass index

index_driors <- sar_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
                                 catch = ex$catch, 
                                 years = ex$year,
                                 sar = sar$sar[sar$stockid == unique(ex$stockid)],
                                 index = ex$b_v_bmsy[!is.na(ex$b_v_bmsy)], 
                                 index_years = ex$year[!is.na(ex$b_v_bmsy)])

index_fit <- fit_sraplus(index_driors, engine = "stan", estimate_shape = TRUE)

plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit, `sar fit` = sar_fit,
             `index fit` = index_fit,
             plotvars = "b_div_bmsy") + 
  geom_point(data = ex, aes(year, b_v_bmsy))
